Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016

### 2019-10-26 response_for_ANZJS_editor.Rmd

---
title: "Response to the Referees"
author: "Jeremy Forbes, Di Cook, Rob Hyndman"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

@@

Dear Dr Holland,

Thank you for your helpful comments on our paper titled 'Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016'.

We have revised our paper to address the issues raised. Below are detailed summaries of the changes made in response to the points raised by yourself, the Associate Editor and by the anonymous reviewers.

## Editor's Comments

@@

1. Address reviewer 1's question about why spatial imputation is required given that the census is reported at a low level of aggregation (lines 150-151) that could presumably be clumped to give a good approximation to electorate boundaries.

*We now use Census data at the lowest level of aggregation available --- the geographic region type of 'Statistical Area 1' --- to approximate the electoral boundaries (see Section 2.2). Some spatial imputation is still required as the SA1 boundaries do not exactly match the electorate boundaries. Even with this change in imputation method, our results are effectively the same as the previous submission.*

@@

2. I note that you do mention the ecological fallacy (lines 285-286) but reviewer 1's reaction suggests that you should expand your remarks here.

*We have expanded our remarks and moved them to the introduction (Section 1) to give them more prominence. We are explicit in stating that "Insights are being drawn at the electorate level and cannot be inferred for another disaggregate level (in particular, drivers of individual voter behaviour may vary from what is observed at the electorate level)".*

@@

3. It appears that your approach to variable selection has confused everybody. My understanding of what you have done is that: You begin with 65 variables (are they all the available census data or a subset?). Then you standardise the variables. Then you remove some variables to do with Age Bands and Christianity (how many variables out of the 65 does this leave? Were these variables totally removed or lumped in some way? Christianity still appears in table 2 so they obviously weren't all removed)

*We have now explained our approach to variable selection in more detail and made some additional changes to to provide greater clarity. Previously we started with 65 variables, but many of these variables were removed in the variable selection process because they provided additional granularity that we did not want to pursue. We have now refined the subset of variables considered for analysis in this paper to 50 Census variables. Census data is provided as a head count for categories for many different questions, and there are thousands of variables that could be constructed from the raw data. These 50 variables were chosen because they appear in previous studies of Australian elections (those mentioned in the Introduction). We hope this is now better explained.*

4. ...Then you use the PCA to identify variables that covary. If variables appear together in a PC with loading higher than 0.15 they are combined together into a factor, e.g. a new variable Income replaces two previously existing variables. (In the description of this it is unclear how many variables are being combined in each factor, the ‘ands’ and commas make things tricky, .e.g is ‘renting and government housing’ one variable or two? Do the factors use the loading weights or just the signs?)

*We have made an effort to clarify this process in Section 3.2, which now details each of the variables being combined into each factor. For consistency, only commas are used to separate the variable descriptions.*

*The factors are calculated using only the signs, not the loading weights. A total of 23 variables are combined into 6 different factors.*

@@

5. The steps above get you from 65 variables down to 30 (I can count 10 variables that  get accounted for via the creation of factors, but that doesn't get from 65 to 30 so presumably some other variables have been removed? Or is this difference all in the age bands and variants of Christianity? More clarity is required)

*This has been reconciled in Section 3.2. We begin with 50 variables. Six factors are created using a total of 23 variables. One of the age brackets (populated aged 55 and over) is removed, and the other three age brackets are kept (to avoid multicollinearity). This results in a final predictor set of 32 variables. Any variables that are not contained in a factor are not discarded, and are kept in the model as a explanatory variables.*

@@

6. As you note in the paper, it seems likely that there are still issues with multi-collinearity. Figs 6 and 7 show the variables with significant effects, do the results change much if you fit a model with just these 13 variables instead of the 30?

*This was a very helpful suggestion to explore, and as such an additional section on robustness has been included in the revised paper (Section 4.4). When we fit a model using just the variables with significant effects, we find that the results are robust. Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).*

@@

*A third and final check is the visual exploration of different variable projections using a tour (for each election). No definitive signs of multicollinearity are observed.*

7. Other minor points

  * line 120,  each of the
  * line 129, is it really the boundaries that are the target /source zones or the areas defined by these boundaries?
  * line 241, must be modelled
  * line 261-266, is there anything in the fact that the spatial coefficient is significant only in the years where no imputation was done?

*The first three minor issues have been corrected. For the last issue, the spatial coefficient is now significant in 2001, 2004 and 2016, which do not correspond only to years with no imputation.*

## Associate Editor's Comments

1. The main thing I found confusing about this paper was the variable selection procedure. It seems to be heading in the direction of four PCs, then it veers into six factors before ending with 30 variables. This all requires greater clarity and justification.

*We have reworked Section 3.2 to better explain this process.*

*We begin with 50 variables. Many of these variables represent similar information and where appropriate, should be grouped together. PCA is used to determine which variables covary, and hence decide which varibles should be combined into a factor. Each of the first four PCs are examined to determine which variables have large loadings. If there is an intuitive reason as to why a subset of variables with large loadings (in a given PC) should be grouped together, then they are combined into a factor using a weighted sum. The weights are simply +1 is the loading in that PC is positive, and -1 if it is negative.*

*Six factors are created using a total of 23 variables. One of the age brackets (populated aged 55 and over) is removed, and the other three age brackets are kept (to avoid multicollinearity). This results in a final predictor set of 32 variables. Any variables that are not contained in a factor are not discarded, and are kept in the model as a explanatory variables.*

2. The authors opt for “visualization” of the fixed effects rather than the usual table of estimated parameters. Perhaps this is a good choice given that there are so many. But then spatial parameters are exhibited in a huge table. This seems inconsistent.

@@

*Partial residual plots are presented to give the reader a visual feel for the direction, size and significance of an estimated effect, as well as the distribution of the data points (for that covariate). Whilst there is duplication of information here, we think that the partial residual plots play an important role in illustrating the effect of a given variable across elections, and that the estimated model parameters should be included for completeness.*

3. I found the writing unclear in a few places. What are the “time-varying restrictions” on a variable (line228)?

*We have revised our explanation relating to the model specification (Section 3.3). Each election is modelled as a separate cross section, which can be considered a special case of a longitudinal model where all coefficients are time-varying and heteroskedasticity is time-varying. A time-varying coefficient is a coefficient that a different estimate for each time period $t=1,...,T$. We have removed the note about "time-varying restrictions".*

4. And the description of a partial residual (line 276) doesn’t make sense to me.

*Partial residuals, for a variable X1, are the residuals from the fitted model with the estimated effect of X1 added to each. Partial residual plots show the direction, size and significance of an estimated effect, as well as any deviations from linearity (see Section 4.3).*

@@

5. The authors recognise that there is likely to be collinearity among 65 Census variables – why those; we are never told?

*The (now) 50 Census variables are chosen because they appear in previous studies of Australian elections (those mentioned in the Introduction), and we believe that they capture key characteristics of an electorate - age, income, employment, religion, birthplace and relationships.*

*Note that we have refined the subset of variables considered for analysis in this paper to 50 Census variables. Previously we started with 65 variables, but many of these variables were removed in the variable selection process because they provided additional granularity that we did not want to pursue.*

6. And so, as others do, they did a pca of the data matrix, reducing it to four components – we are not told what rotation was employed to achieve simple structure. We are not given the results of this analysis, merely a naming of the six factors identified – how do you get six ‘factors’ from four components?

*We have reworked Section 3.2 to better explain this process.*

*Our approach is to combine the six election datasets together into a single dataset of 900 observations. We computed the principal components on this. The first four components are chosen because the amount of variance explained drops off after the fourth PC. We then use these four PCs to determine which variable covary, and hence decide which varibles should be combined into a factor.*

*Each of the first four PCs are examined to determine which variables have large loadings. If there is an intuitive reason as to why a subset of variables with large loadings (in a given PC) should be grouped together, then they are combined into a factor using a weighted sum. The weights are simply +1 is the loading in that PC is positive, and -1 if it is negative. Hence, the factors are not the principal components themselves.*

7. My expectation was that they would then fit a model using the four components as the independent variables, thus avoiding any collinearity and confounding. But they didn’t. Instead they selected 30 of the 65 variables – which must still be collinear; the resulting equation in Table 2 clearly implies this given the large number of (a) insignificant relations (normal with collinearity – see a recent paper in Quality and Quantity), (b) unreasonably large coefficients (-8.07 for de facto, for example), and (c) very large standard errors for many of the variables.

*In response to this suggestion, we have now added an assessment of multicollinearity within the robustness section of the revised paper (Section 4.5). Three robustness checks are conducted as follows:*

- *We fit a model using just the variables with significant effects, we find that the results do not change much. Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).*

- *We also checked the effect of omitting a variable that is contained in the ten largest pairwise correlations. For each pair, a model for each election is refit omitting one of the two variables. It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the $95/%$ confidence interval from the full model.*

- *A third and final check is the visual exploration of different variable projections using a tour (for each election). No definitive signs of multicollinearity are observed.*

8. The authors fit a spatial autoregressive model, on the assumption that neighbouring places are more likely to be similar than distant places (the ‘first law of geography’). This is only a significant finding at two of the elections and is wrongly interpreted: it does not necessarily show that then ‘the political climate of an electorate appears to be affected by the attitudes of its neighbours’ but, much more likely, that neighbouring places share characteristics in common that affect voting behaviour but are not captured by the independent variables included in the model. Spatial autocorrelation may reflect spatial diffusion processes – but in which direction? – but is much more likely to reflect an under-specified model.

*In retrospect, we agree that this is perhaps a naïve conclusion. It is possible that unobserved variables are being captured in the spatial effect, which is the interpretation that we have now included. We have removed the old interpretation from the revised paper. Additionally, on revision of the imputation method, we find that the spatial effect is significant in 2004 as well as 2001 and 2016.*

## Reviewer 2's comments

1. It seems very strange to assume that population is uniform within electorates (line 133) so that weighting is proportional to area, and not to the number of people in an area of intersection. An electorate on the urban fringe may be strongly affected by loss of a relatively small but highly populated area but little affected by incorporating a large rural area. A simple indirect way to get at this might be to address the question: Would the results change much if A_{s,t} was log area, for each s,t ?

@@

*We have revised our imputation method to avoid this assumption. We now use Census data at the lowest level of aggregation available - the geographic region type of 'Statistical Area 1' (SA1) - to obtain the approximation of electoral boundaries (see Section 2.2). There are on average approximately 400 people in each SA1. This approach will successfully capture any densely populated areas on the fringe of an electorate. Our original method was chosen because we believed that the imputation method would not materially affect results. With this change in imputation method, our results are effectively the same as the previous submission.*

2. It seems to me that the use of Principal Components analysis is unclear. (a) PC appears to have been used as a model selection criterion, with the variables related to the first four PCs included into the model in Table 2. However typically when principal components are computed they only involve the explanatory variables, not the response, and so there is no guarantee that all the first four PCs (and only them) will be related to the response. For example using the eechidna data for 2016 I found the fifth PC was extremely significant. But this is a minor point.

*We have used principal components to inform which variables should be combined into factors, rather than use them directly as explanatory variables. This is done as a step to address any (near) multicollinearity in the explanatory variables. Any variables that are not used in creating a factor still remain in the model. Of course, there is no guarantee about any of these factors being related to the response. We have now explained our variable selection procedure in more detail in the revised paper (Section 3.2).*

3. The variables which load highly on a PC are likely to be highly correlated, and so if these high-loading variables are used together in the regression model we may again have multicollinearity. I found this was the case when applying the author’s regression model to the 2016 data. This may be why several of the predictors in Table 2 are never related significantly related to the response in any of the years.

*By taking a subset of variables with high loadings on a principal component and combining them into a single variable, we are trying to address potential multicollinearity. Two factors that have been created from the same PC will be correlated, which could mean that (near) multicollinearity persists. However, we have conducted three robustness tests for multicollinearity (see Section 4.5), and no conclusive signs of multicollinearity were observed.*

@@

*The first check involved fitting a model using just the variables with significant effects. We find that the estimated coefficients in the reduced model (containing only the subset significant variables) all lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).*

*We also checked the effect of omitting a variable that is contained in the ten largest pairwise correlations. For each pair, a model for each election is refit omitting one of the two variables. It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the $95/%$ confidence interval from the full model.*

*A third and final check is the visual exploration of different variable projections using a tour (for each election). No definitive signs of multicollinearity are observed.*

4. Conversely there could be a variable which is independent of the other variables (so representing 1 / 65th of the explanatory information) but by dint of that fact not be included in the first four PCs. In short, the narrative leaves me unconvinced that all the variables included are necessary or that all the relevant variables are included in the regression.

@@

*Any variables that are not used in creating a factor still remain in the model. The four PCs are used ONLY to inform which variables are correlated and should be combined into a factor. For example, median household, personal and family incomes all have large positive loadings in the first principal component, so we combine them into a factor called `Incomes`. Any variable that does not have a large loading in any of the four PCs are not discarded - they remain in the model. It is of course possible that there exist variables beyond what we have considered that could affect two party preference. However, we believe that the variables we have used cover the core demographic components of an electorate - age, income, employment, religion, birthplace and relationships.*

5. I presume the “factors” referred to lines 210-220 were created as the first principal component of the constituent raw variables?*

*In our revised submission, we have made an effort to clarify what we have defined a "factor" to be (Section 3.2). A "factor" is created when there is evidence to suggest that a group of variables are explaning similar information, and if they were all included in the model as separate variables, there would be severe issues with multicollinearity. To inform which variables should be grouped, we use the first four PCs from PCA. In a single PC, if there are is a subset of variables with large (positive or negative) loadings that intuitively could represent similar information, we combine them into a factor using a weighted sum, where the weights are +1 or -1 depending on the sign of the loading in the that PC. This means that many factors can be created from a single PC, and that any variables that do not appear in a factor are still included in the model as stand-alone variables.*

6. The description of the variables included in “Education” is not clear.

@@

*We have provided more details on the variables that are combined into each of the factors, including `Education`. `Education` is a weighted sum of seven variables that represent: high school completions, undergraduate and postgraduate degrees, proportion of employed people working as professionals, proportion of jobs in finance, proportion of workers who are laborers, proportion of workers who work as a tradesperson, diploma and certificate qualifications.*

7. I could not get the variable ‘OtherLanguageHome’ in 2016 to be significant, but it also had a high VIF with other explanatory variables. Please check this.

### 2019-11-10 anzjs_elections.Rmd
minor typo fixes

Socio-demographic variables were derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income. There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016. The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information. The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections. From the available Census information aggregated at the electorate level, 50 socio-demographic variables were defined for each of the electorates to be used in the analysis. These variables include information relating to electoral age distributions, income, education qualifications, employment industries and job types, religion, birthplace, household characteristics and relationships.

@@

For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election. This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of neighbouring Censuses, thereby addressing the spatial aspect. Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.

@@

An identical model specification is used for each of the six elections, with each election modelled separately. Separate models are preferred to a single model because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographic region. Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation (of voting information). The separate models also allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects. This can be considered a special case of a longitudinal model where all coefficients are time-varying and heteroskedasticity is time-varying.

@@

```{r prepplot, include = FALSE}
p1 <- grid_visreg("Extractive", top = TRUE)
p2 <- grid_visreg("ManagerAdmin")
p3 <- grid_visreg("Transformative")
p4 <- grid_visreg("SocialServ")
p5 <- grid_visreg("DiffAddress")
p6 <- grid_visreg("Incomes")
p7 <- grid_visreg("OtherLanguage")
p8 <- grid_visreg("Pop_00_19")

@@

Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect. Of these areas, only South-Eastern European populations significantly affect party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \@ref(fig:partresplot2)). Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguage` variable. Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 row 7, Figure \@ref(fig:partresplot1)). Furthermore, none of the variables relating to religious beliefs aside from Christianity have a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations). The association between Christian populations (`Christianity`) and the Liberal party steadily increases over the years, becoming positive and significant in 2016.

@@

It is often hypothesized that states have a systematic bias towards one of the two major parties. Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) reveal that the data reflects this to only a limited extent. Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the South Australia and the Northern Territory tend towards voting Liberal. However, there are relatively few electorates in each of these states (five, two, eleven and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.

@@

The authors would like to sincerely thank the editor and associate editor of the Australian & New Zealand Journal of Statistics and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of the manuscript. Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.

>>>>

Socio-demographic variables are derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income. There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016. The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information. The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections. From the available Census information aggregated at the electorate level, 50 socio-demographic variables are defined for each of the electorates to be used in the analysis. These variables include information relating to electoral age distributions, income, education qualifications, employment industries and job types, religion, birthplace, household characteristics and relationships.

@@

For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election. This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of the neighbouring Censuses, thereby addressing the spatial aspect. Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.

@@

An identical model specification is used for each of the six elections, with each election modelled separately. Separate models are preferred to a single model because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographic region. Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation (of voting information). The separate models also allow the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects. This can be considered a special case of a longitudinal model where all coefficients are time-varying and heteroskedasticity is time-varying.

@@

```{r prepplot, include = FALSE, out.height="100%"}
p1 <- grid_visreg("Extractive", top = TRUE)
p2 <- grid_visreg("ManagerAdmin")
p3 <- grid_visreg("Transformative")
p4 <- grid_visreg("SocialServ")
p5 <- grid_visreg("DiffAddress")
p6 <- grid_visreg("Incomes")
p7 <- grid_visreg("OtherLanguage")
p8 <- grid_visreg("Pop_00_19")

@@

Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect. Of these areas, only South-Eastern European populations significantly affect party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \@ref(fig:partresplot2)). Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguage` variable. Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 (see row 7, Figure \@ref(fig:partresplot1)). Furthermore, none of the variables relating to religious beliefs aside from Christianity have a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations). The association between Christian populations (`Christianity`) and the Liberal party steadily increases over the years, becoming positive and significant in 2016.

@@

It is often hypothesized that states have a systematic bias towards one of the two major parties. Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) show that the data reflects this to only a limited extent. Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the South Australia and the Northern Territory tend towards voting Liberal. However, there are relatively few electorates in each of these states (five, two, eleven and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.

@@

The authors would like to sincerely thank the editor and associate editor of the Australian & New Zealand Journal of Statistics and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of this manuscript. Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.
